### 2017-09-26 paper/article_0626.tex
adds some more improvements to the introduction

\section{Introduction}
Modeling for disease surveillance data to explore patterns of infectious diseases has had a long history in public health. Infectious diseases can cause high economic and medical costs due to morbidity and mortality. In recent decades, the annual number of global deaths caused by infections has leveled off at approximate 15 million and may remain at this level for the next three decades \cite{DyeC, WHOM}. In order for such an enormous health burden to be reduced, preventing and controlling infectious diseases becomes extraordinarily important, which depends on how much we know about the course of disease transmission. For zoonotic diseases, where transmission to humans may be via food, water, direct contact with animals, or through environmental contamination, knowledge of the potential sources and pathways of infection is key in improving understanding of the burden of disease. Quantifying the burden of illness to each pathway and source, provides valuable information to health professionals making decisions on public health policy for disease control.

Over the last few decades there have been many models developed for the analysis of zoonotic diseases. Epidemiological models make use of the incidence of cases and geographic data to capture the main characteristics of disease spread. For example, the Susceptible-Infectious-Recovered model \cite{Kerm} classifies populations into a susceptible, infective and recovered status enabling an estimation of the number of individuals who are likely to be infected. Also, models with a combination of spatial and temporal data enable the prediction of the time frame and areas where a disease outbreak is likely to occur \cite{HeldL, Hoehl, Simo}. However zoonotic transmission from animals to humans may be complex, involving many sources and exposures linked by differing pathways. For instance, infected wild birds may contaminate environmental water and cause the spread to water users, either humans or non-humans \cite{Wagen}. Tracing the source of infection is thus crucial to increasing the ability to implement risk management and intervention \cite{Wilso, Morel}. In terms of statistical analysis, it means that modelling zoonoses requires an advanced approach with the focus changed from just epidemiology to a combination of epidemiology, evolutionary genetics and biology \cite{Muell}. Hence, some source attribution models have been proposed to estimate the number of cases attributable to different sources by utilizing epidemiological information and the association with genotypes found in humans and sources \cite{vanP, Hald, MullA}.

@@

The genetic information used in such integrated models is typically at the level of genotyping that groups closely related organisms together to capture broader changes in DNA occurring in the transmitted pathogens \cite{Cotta}. Typically molecular typing methods, such as multilocus sequence typing (MLST) \cite{Dingl, Coll} might be used, which utilises nucleotide sequences of internal fragments of a small set of housekeeping genes. Such sequences have sufficient variation to distinguish differing pathogen lineages, while being relatively stable within lineages. Each unique nucleotide sequence (allele) at each housekeeping gene (loci) is assigned a number, and the set of numbers across all loci (the allelic profile) is then taken as the genotype, with is usually also assigned a number.

For the pathogen \textit{Campylobacter}, which causes campylobacteriosis, a worldwide gastrointestinal disease in humans. For \textit{Campylobacter}, the MLST scheme consists of housekeeping genes \textit{asp}A (aspartase A), \textit{gln}A (glutamine synthetase), \textit{glt}A (citrate synthase), \textit{gly}A (serine hydroxymethyltransferase), \textit{pgm} (phosphoglucomutase), \textit{tkt} (transketolase), and \textit{unc}A (ATP synthase $\alpha$ subunit). An illustrative example of a part of MLST data for \emph{Campylobacter jejuni} is presented in Table~(\ref{tab0}). It shows that the genotypes ST-45 and ST-474 have different allelic combinations across the seven loci. The different allelic profiles enable comparison of gene similarities or dissimilarities so that an association between sources and infected cases can be made.

>>>>

\section{Introduction}
Modeling of disease surveillance data to explore patterns of infectious diseases has had a long history in public health. Infectious diseases can cause high economic and medical costs due to morbidity and mortality. In recent decades, the annual number of global deaths caused by infections has leveled off at approximate 15 million and may remain at this level for the next three decades \cite{DyeC, WHOM}. In order for such an enormous health burden to be reduced, preventing and controlling infectious diseases becomes extraordinarily important, which depends on how much we know about the course of disease transmission. For zoonotic diseases, where transmission to humans from animal reservoirs may be via food, water, through environmental contamination or via direct contact with animals, knowledge of the potential sources and pathways of infection is key in improving understanding of the burden of disease. Quantifying the burden of illness to each pathway and source, provides valuable information to health professionals making decisions on public health policy for disease control.

Over the last few decades there have been many models developed for the analysis of zoonotic diseases. Epidemiological models make use of the incidence of cases and geographic data to capture the main characteristics of disease spread. For example, the Susceptible-Infectious-Recovered model \cite{Kerm} classifies populations into a susceptible, infective and recovered status enabling an estimation of the number of individuals who are likely to be infected. Also, models with a combination of spatial and temporal data enable the prediction of the time frame and areas where a disease outbreak is likely to occur \cite{HeldL, Hoehl, Simo}. However zoonotic transmission from animals to humans may be complex, involving many sources and exposures linked by differing pathways. For instance, infected wild birds may contaminate environmental water and cause disease spread to water users, either humans or other animals \cite{Wagen}. Tracing the source of infection is thus crucial to increasing the ability to implement risk management and intervention \cite{Wilso, Morel}. In terms of statistical analysis, modelling zoonoses requires an advanced approach with the focus changed from just epidemiology to a combination of epidemiology, evolutionary genetics and biology \cite{Muell}. Hence, some source attribution models have been proposed to estimate the number of cases attributable to different sources by utilizing epidemiological information and the association with genotypes found in humans and sources \cite{vanP, Hald, MullA}.

@@

The genetic information used in such integrated models is typically through genotyping that groups closely related organisms together to capture broader changes in DNA occurring in the transmitted pathogens \cite{Cotta}. Typically molecular typing methods, such as multilocus sequence typing (MLST) \cite{Dingl, Coll} might be used, which utilises nucleotide sequences of internal fragments of a small set of housekeeping genes. Such sequences have sufficient variation to distinguish differing pathogen lineages, while being relatively stable within lineages. Each unique nucleotide sequence (allele) at each housekeeping gene (loci) is assigned a number, and the set of numbers across all loci (the allelic profile) is then taken as the genotype, with is usually also assigned a number.

For the pathogen \textit{Campylobacter} which causes campylobacteriosis, a worldwide gastrointestinal disease in humans, the MLST scheme consists of housekeeping genes \textit{asp}A (aspartase A), \textit{gln}A (glutamine synthetase), \textit{glt}A (citrate synthase), \textit{gly}A (serine hydroxymethyltransferase), \textit{pgm} (phosphoglucomutase), \textit{tkt} (transketolase), and \textit{unc}A (ATP synthase $\alpha$ subunit). An illustrative example of MLST data for \emph{Campylobacter jejuni} is presented in Table~(\ref{tab0}). It shows that the genotypes ST-45 and ST-474 have different allelic combinations across the seven loci. The different allelic profiles enable comparison of gene similarities or dissimilarities so that an association between sources and infected cases can be made, by comparing the distribution of genotypes from human cases with those from potential reservoirs.
